---
title: "Comparison with Base R's lm()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparison with Base R's lm()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette demonstrates how to use the `biost625linreg` package to fit linear regression models, compares it to base R's `lm()`, and verifies its correctness.

# Data Setup

We will use the `iris` and `mtcars` datasets to illustrate.

```{r setup, echo=TRUE}
# Loading required packages
library(biost625linreg) # My package

library(bench) # For benchmarking

# Loading the data
data(iris)
data(mtcars)
```

# Demonstration of package using the `iris` dataset

### Fitting a linear model
```{r}
# Fit a linear regression model using biost625linreg with the below code:
my_model <- linear_regression_fit(Sepal.Length ~ Petal.Length + Petal.Width, iris)

# View the coefficients using:
print(my_model$coefficients)

print(head(my_model$residuals))

print(my_model$r_squared)
```

The `linear_regression_fit` function fits a linear regression model and returns the coefficients, residuals, and R-squared value. The coefficients help with making predictions, residuals are beneficial for further analysis like creating a qq plot to understand the dataset better, and R-squared value gives us an idea of how well the model is fitting the data.


### Extracting predictions out of fit model
```{r}
# Extract predictions using:
my_predictions <- linear_regression_predict(my_model, iris)

# Display the predictions using:
head(my_predictions)
```

The `linear_regression_predict` function allows us to generate predictions for the `Sepal.Length` variable based on the fitted model.

### Evaluating the model
```{r}
my_r_squared <- my_model$r_squared
print(my_r_squared)
```

This means that the model is fitting fairly well.
```{r}
my_rmse <- rmse(iris$Sepal.Length, my_predictions)
print(my_rmse)
```

The `rmse` function computes the Root Mean Squared Error, providing insight into model accuracy. A low RMSE score in this case further indicates that the model is able to explain the variance of the dataset well.

# Comparison with Base LM package using the `mtcars` dataset
```{r}
# Fit using biost625linreg
my_model <- linear_regression_fit(mpg ~ wt + hp, mtcars)

# Fit using base R
base_model <- lm(mpg ~ wt + hp, data = mtcars)
```

```{r}
# Coefficients from my package
my_model$coefficients

# Coefficients from base R
coef(base_model)

all.equal(unname(my_model$coefficients), unname(coef(base_model)))
```
Looking at the above output we can see that my linear regression module produces the exact same coefficients as the base lm module.
```{r}
# Predictions using my package
my_predictions <- linear_regression_predict(my_model, mtcars)
print("My LM's Predictions:")
print(my_predictions)

# Predictions using base R
base_predictions <- predict(base_model, newdata = mtcars)
print("Base LM's Predictions:")
print(base_predictions)

all.equal( unname(my_predictions), unname(base_predictions))
```
Looking at the above output we can see that my linear regression module produces the exact same prediction for each row as the base lm module.

```{r}
# Computing R-squared using my package
my_r_squared <- my_model$r_squared
print("My LM's R-squared:")
print(my_r_squared)

# Computing R-squared using base R
base_r_squared <- summary(base_model)$r.squared
print("Base LM's R-squared:")
print(base_r_squared)

all.equal(my_r_squared, base_r_squared)
```
The r-squared values for both the packages are the same, which indicates that the package is working as expected and is reliable.

```{r}
# Computing RMSE using my package
my_rmse <- rmse(mtcars$mpg, my_predictions)
print("My LM's RMSE:")
print(my_rmse)

# Computing RMSE manually using base R
base_rmse <- sqrt(mean((mtcars$mpg - base_predictions)^2))
print("Base LM's RMSE:")
print(base_rmse)

all.equal(my_rmse, base_rmse)
```
The RMSE also seems to be identical between my LM package and the base LM package. 

```{r}
bench::mark(
  biost625linreg = linear_regression_fit(mpg ~ wt + hp, mtcars)$coefficients,
  base_r = coef(lm(mpg ~ wt + hp, data = mtcars))
)
```
The benchmarks show that the fit function from my module is faster than the base lm package. The minimum, median and total runtimes are lower for my function when compared to the lm module from the stat package. 

```{r}
bench::mark(
  biost625linreg = linear_regression_predict(my_model, mtcars),
  base_r = predict(base_model, newdata = mtcars)
)
```

It is a similar story for the predict function as my function outperforms the base function in terms of speed.
